Corotating Clumps Around Adolescent Low-Mass Stars: Four Years of TESS

### 2023-07-05 paper/ms.tex
update ms

% primary authors
\author{Author list and order not final!}
\author[0000-0002-0514-5538]{Luke~G.~Bouma}
\altaffiliation{51 Pegasi b Fellow}
\affiliation{\caltech}

@@

A few competing explanations for what causes the complex quasiperiodic
variability are shown in Figure~\ref{fig:f1}, along with some
representative light curves.  All young M-dwarfs are spotted, which
produces flux variations over characteristic timescales of the
rotation period, $P_{\rm rot}$, and its half-harmonic, $0.5\,P_{\rm
rot}$.  The observed dips occur over durations as short as
$0.05\,P_{\rm rot}$; a ``starspot-only'' scenario can be discarded for
any object with sufficiently sharp dips
\citep{2017AJ....153..152S,2021MNRAS.500.1366K}.   The timescales and
amplitudes of the flux variations instead require sharp geometries
with material extrinsic to the stellar surface
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.  The
``clump'' scenario invokes opaque dust clumps orbiting near the
Keplerian corotation radius, which periodically eclipse the star
\citep{2017AJ....153..152S,2023MNRAS.518.4734S}.  The ``prominence''
scenario invokes long-lived condensations of cool and dense
marginally-ionized gas, embedded within the hotter corona, that would
be centrifugally supported near corotation
\citep{1989MNRAS.238..657C,2019MNRAS.482.2853J,2022MNRAS.514.5465W}.
Such structures are analogous to quiescent prominences and filaments
observed in the solar corona \citep[see e.g.][]{2015ASSL..415.....V},
though at much larger relative distances from the stellar surface.  A
final possibility that has been suggested is one of a ``screen'', in
which the inner wall of a quiescent circumstellar disk blocks a
portion of the stellar surface to produce sudden dips whenever spots
come into view \citep{2019ApJ...876..127Z}.

@@

One way to help rule between the mechanisms is to therefore find
bright and nearby CQVs, since these objects will be the most amenable
to detailed photometric and spectroscopic analyses.  To do this, in
this work, we use 120-second cadence data acquired by TESS between
July~2018 and Sep~2022 (Sectors 1-55; Cycles 1-4).  We present our
search methods in Section~\ref{sec:methods}, and the properties of the
resulting CQV catalog in Section~\ref{sec:results}.  The evolution of
many CQVs over a two-year baseline is described in
Section~\ref{sec:evoln}, including a deep-dive into LP 12-502.
Implications are discussed in Section~\ref{sec:discussion}, and we
conclude in Section~\ref{sec:conclusion}.

A word on nomenclature.  CQVs have been called ``transient and
persistent flux dips'', ``scallop shells'', ``batwings'',
\citep{2017AJ....153..152S} ``complex rotators'',
\citep{2019ApJ...876..127Z,2022AJ....163..144G,2023ApJ...945..114P}
and ``complex periodic variables'' \citep{2023MNRAS.518.2921K}.  
The CQVs should generally not be conflated with ``dippers'', which are classical T
Tauri stars with infrared excesses, and drastically and 
at best marginally periodic variability linked
to obscuring inner disk structures and accretion hot spots
\citep{2014AJ....147...82C,2021ApJ...908...16R}.  
With that said, a few similarities between CQVs and dippers do exist
(see Sections~\ref{subsec:irexcess} and~\ref{subsec:discdippers}).
At the risk of
introducing yet another standard, we prefer a nomenclature that
reflects how, when observed over timescales of more than tens of
cycles, CQVs are almost but not exactly periodic.  They are
irregularly periodic, or for short, quasiperiodic.  While the
three-type classification scheme proposed by
\citet{2017AJ....153..152S} may indeed provide some helpful visual
distinctions amongst CQV sub-classes, it seems likely that they are
all explained by a single underlying phenomenon, and so we opt to
refer to them by a single empirically descriptive name.

@@

We determined the stellar effective temperature and radii through SED
fitting; we then estimated the masses by interpolating against the
PARSEC v1.2S models \citep{2012MNRAS.427..127B,2014MNRAS.444.2525C}.

>>>>

% primary authors
\author{Author list \& order not final!}
\author[0000-0002-0514-5538]{Luke~G.~Bouma}
\altaffiliation{51 Pegasi b Fellow}
\affiliation{\caltech}

@@

A few competing explanations for what causes the complex quasiperiodic
variability are shown in Figure~\ref{fig:f1}, along with some
representative light curves.  All young M-dwarfs are spotted, which
produces flux variations over characteristic timescales of the
rotation period, $P_{\rm rot}$, and its half-harmonic, $0.5\,P_{\rm
rot}$.  The observed dips occur over durations as short as
$0.05\,P_{\rm rot}$; a ``starspot-only'' scenario can be discarded for
any object with sufficiently sharp dips
\citep{2017AJ....153..152S,2021MNRAS.500.1366K}.   The timescales and
amplitudes of the flux variations instead require sharp geometries
with material extrinsic to the stellar surface
\citep[e.g.][]{2017AJ....153..152S,2022AJ....163..144G}.  The
``clump'' scenario invokes opaque dust clumps orbiting near the
Keplerian corotation radius, which periodically eclipse the star
\citep{2017AJ....153..152S,2023MNRAS.518.4734S}.  The ``prominence''
scenario invokes long-lived condensations of cool and dense
marginally-ionized gas, embedded within the hotter corona, that would
be centrifugally supported near corotation
\citep{1989MNRAS.238..657C,2019MNRAS.482.2853J,2022MNRAS.514.5465W}.
Such structures are analogous to quiescent prominences and filaments
seen in the solar corona \citep[see e.g.][]{2015ASSL..415.....V},
though at much larger relative distances from the stellar surface.  A
final possibility that has been suggested is one of a ``screen'', in
which the inner wall of a quiescent circumstellar disk blocks a
portion of the stellar surface to produce sudden dips whenever spots
come into view \citep{2019ApJ...876..127Z}.

@@

One way to help rule between the mechanisms is to find bright and
nearby CQVs, since these objects will be the most amenable to detailed
photometric and spectroscopic analyses.  To do this, in this work we
use 120-second cadence data acquired by TESS between July~2018 and
Sep~2022 (Sectors 1-55; Cycles 1-4).  We present our search methods in
Section~\ref{sec:methods}, and the properties of the resulting CQV
catalog in Section~\ref{sec:results}.  The evolution of many CQVs over
a two-year baseline is described in Section~\ref{sec:evoln}, including
a deep-dive into LP 12-502.  Implications are discussed in
Section~\ref{sec:discussion}, and we conclude in
Section~\ref{sec:conclusion}.

A word on nomenclature.  CQVs have been called ``transient and
persistent flux dips'', ``scallop shells'', ``batwings'',
\citep{2017AJ....153..152S} ``complex rotators'',
\citep{2019ApJ...876..127Z,2022AJ....163..144G,2023ApJ...945..114P}
and ``complex periodic variables'' \citep{2023MNRAS.518.2921K}.  The
CQVs should not be conflated with ``dippers'', which are classical T
Tauri stars with infrared excesses, and large-amplitude variability
linked to obscuring inner disk structures and accretion hot spots
\citep{2014AJ....147...82C,2021ApJ...908...16R}.  With that said, a
few similarities between CQVs and dippers do exist (see
Sections~\ref{subsec:irexcess} and~\ref{subsec:discdippers}).  At the
risk of introducing yet another standard, we hope to introduce a
nomenclature that reflects how, when observed over timescales of more
than tens of cycles, CQVs are almost but not exactly periodic.  They
are quasiperiodic.  While the three-type classification scheme
proposed by \citet{2017AJ....153..152S} may indeed provide some
helpful visual distinctions amongst CQV sub-classes, it seems likely
that they are all explained by a single underlying phenomenon, and so
we opt to refer to them by a single empirically descriptive name.

@@

We determined the stellar effective temperature and radii by 
fitting the broadband spectral energy distributions (SEDs); we then estimated the masses by interpolating against the
sizes, temperatures, and ages against the
PARSEC v1.2S models \citep{2012MNRAS.427..127B,2014MNRAS.444.2525C}.
